Qualifying Report

### 2019-01-16 Source/QualifyingReport/QualifyingReport.tex
QR adjustments for Brie

In terms of standardised cognitive tests, there are two main aims. Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and, does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility). This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their perfomance in terms of both diagnostic and prognostic utility.

@@

The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specfically looking at detecting a characterizing very mild dementia \cite{Randolph1998}. The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests. They met a number of design goals for this new battery of tests that addressed these shortcomings. The RBANS consists of a number of sub-tests across five distinct domains.

@@

\subsubsection{Digit Span Test}
The Digit Span Test is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers. A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order. The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
% Muangpaisan - Digit span and verbal fluency tests in patients with MCI and normal subjects in Thai community.
MCI patients had significantly lower digit span score, in both normal and reverse order versions of this test. Research has shown that Digits Backwards can, to some degree, predict a diagnosis of MCI. They found that Age, Gender and Education have an impact on the performance of the tests.
Emrani et al found no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups. Notably, there was an attenuated recency effect in those with mixed MCI. \cite{Emrani2018}. Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients. \cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test. No differences were found between forward and backward conditions in any of the groups. However, available tests may not detect subtle impairments \cite{Kessels2015}.
\subsubsection{Rey Auditory-Verbal learning test}
Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processess including short-term auditory-verbal memory, retention of information. Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat. Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes.
Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD. Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints. Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores. They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting. Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}.
% Schoenberg - Test performance and classification statistics for the Rey Auditory Verbal Learning Test in selected clinical samples.
Study found that RAVLT to best distguish patients suspected of Alzheimer's disease from the psychiatric comparison group.
\subsubsection{Digit Symbol Substitution Test}
The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol. Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times. The participant is asked to fill in the corresponding symbol for each number. The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
% Rosano - Digit symbol substitution test and future clinical and subclinical disorders of cognition, mobility and mood in older adults.
Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders. Associations were stronger for incident clinical disorders in cognition. Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood. While in part attributable to vascular brain disease, other potentially modifiable contributors may be present. Further studying the causes of psychomoto slowing with ageing might provide novel insights into age-related brain disorders.
% Pascoe - The symbol-digit modalities test in MCI: Evidence from Parkinson's Disease Patients.
Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants. PD-MCI pariticipants achieved significantly lower scores than other groups in the DSST task.  
\subsubsection{Corsi Block Tapping Task}

@@

\subsubsection{Verbal Fluency}

@@

\subsubsection{Naming Tests}

@@

\subsubsection{Rey-Osterrieth Complex Figure Task}
% Cite Rey 1941 - L'examen psychologique dans les cas d'encephalopathie traumatique and Osterrieth 1944 - Le test de copie d'une figure complexe.
The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory. The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944), asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition). In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability. The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
% Salvadori - Qualitative Evaluation of the Immediate Copy of the Rey-Osterrieth Complex Figure: Comparison Between Vascular and Degenerative MCI patients.
Patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory. Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
% Teng - Persistence of Neuropsychological Testing Deficits in Mild Cognitive Impairment.

@@

\subsubsection{Hayling Sentence Completion Test}

@@

\subsubsection{Grooved Pegboard Test}

@@

\subsubsection{Visual Search}

@@

\subsubsection{Free Cued Selective Reminding Test}

@@

\subsection{Types of Language Assessment}

@@

\subsubsection{Interviews}

@@

\subsubsection{Picture-related content thematic elements}

@@

\subsubsection{Perseveration}

@@

\subsection{State of literature into Machine Learning and Natural Language processing techniques}
As we can see, diagnosing dementia through analysis has a large background in terms of psychological research. However, a lot of more current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}. Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language. That is to say that we can subdivide NLP processes into different components. This broadly fits with the literature review conducted by Emery \cite{Emery2000}.
\begin{enumerate}
	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
	\item Syntactic: for grouping and sequencing elementds within samples of language (usually sentences).
	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
	\item Pragmatic: for understanding the context and purpose of an utterance.
\end{enumerate}
\subsubsection{Traditional methods of Machine Learning}
% Thomas - Automatic Detection and Rating of AD through the Lexical Analysis of Spontaneous Speech
Thomas et al, detail several lexical approaches to the problem of detecting and rating DAT. The approaches rely on character n-gram techniques but also explore correlation of usage of frequency of different types of speech. Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech. Results - 95\% accuracy dementia vs controls. 70\% accuracy classifying dementia into two categories. 50\% accuracy classifying dementia into four categories. They suggest further exploration of characteristics to classify.
% Bucks et al - Analysis of spontaneous, conversational speech in dementia of the Alzheimer's Type: Evaluation of an objective technique for analyzing lexical performance.
AD patients had higher mean P rate, A rate and V rate scores but lower N rate scores vs normal older controls \cite{Bucks2000}.
One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class. Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls. Their main finding supported n-grams as the most significant predictor. One of the criticisms is the use of picture description tasks and n-grams. Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
\subsubsection{The case for Deep Learning}

@@

@@

\section{Research Aims}
The aim of my research is to find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests. There is a lot of research into the analysis of language as a marker for MCI and Early Dementia \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}. It is this area which is the focus of my research.

@@

Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language. I focus particularly on MCI as this is often a precursor to AD and therefore detection of MCI in an elderly population would allow I plan to explore the use Natural Language Processing and Machine Learning techniques to assist in this process, and focus on MCI specifically. early diagnosis of dementia and/or mild cognitive impairment through language produced naturally rather than language that is produced as part of a formal cognitive assessment. There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia. These studies have identified a number of speech and language differences between healthy controls and those with dementia. Some of these features have been validated in work using machine learning to classify healthy controls from those with dementia. However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'. While this is help to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments.

@@

Machine Learning itself is undergoing a shift in focus at the present time. I will describe this shift using the context of my proposed research. In traditional machine learning, the transcripts from cognitive assessments are used to generate features such as idea density. These features are then used to classify patients. A company in Toronto, Canada has taken this approach and currently uses over 400 acoustic and language features to make this classification and to generate a psuedo MMSE score.
The broad aim of my research is to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease. Another aim of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD. Finally, an additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.

@@

\section{Research Challenges}
\subsection{Data Collection}

@@

Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other. I'm not going to control for memory problems as a potential confound, but does this weaken the research? How do I defend this? \newline

>>>>

In terms of standardised cognitive tests there are two main aims. Does the test distinguish accurately between normal aging, MCI and AD (diagnostic utility) and does the test distinguish between those individuals with MCI who will then go on to develop AD and those individuals with MCI who don't then go on to develop AD (prognostic utility). This section of the literature review outlines a number of different cognitive tests that have been used to measure cognitive impairments as well as their performance in terms of both diagnostic and prognostic utility.

@@

The Repeatable Battery for the Assessment of Neuropsychological Status (RBANS) was originally developed as an assessment tool for dementia, specifically looking at detecting a characterizing very mild dementia \cite{Randolph1998}. The authors felt that there was a shortfall of appropriate measures that were sensitive enough to milder impairments as well as a number of other shortcomings of existing tests. They met a number of design goals for this new battery of tests that addressed these shortcomings. The RBANS consists of a number of sub-tests across five distinct domains.

@@

\subsection{Digit Span Test}
The Digit Span Test (DST) is predominantly used to measure a person's working memory capacity, specifically the capacity used to store and recall numbers. A participant is presented with a series of numbers of fixed length and is asked to recall those numbers in normal or reverse order. The series of numbers gets progressively longer until such time as a participant fails three or more times out of eight presentations.
\par
% Muangpaisan - Digit span and verbal fluency tests in patients with MCI and normal subjects in Thai community.
Research has consistently shown that MCI patients have a significantly lower digit span score in both normal and reverse order versions of this test and a study by Muangpaisan has shown that the reverse order version of this test can, to some degree, predict a diagnosis of MCI. Muangpaisan also found that Age, Gender and Education have an impact on the performance of the tests. Emrani et al found that the DST revealed no differences between specifically amnestic MCI and controls, but could differentiate between Mixed MCI and the other groups with mixed MCI recalling fewer correct responses than other groups. Notably, there was an attenuated recency effect in those with mixed MCI. \cite{Emrani2018}. Kessels identifies that that there are working memory deficits in MCI patients and these worsen with AD patients. \cite{Kessels2011} He also identified that both MCI and AD have impaired performance on all three conditions of the digit span test. No differences were found between forward and backward conditions in any of the groups. However, available tests may not detect subtle impairments \cite{Kessels2015}.
\subsection{Rey Auditory-Verbal learning test}
Rey's Auditory Verbal Learning Test (RAVLT+) looks a wide range of neuropsychological processess including short-term auditory-verbal memory, retention of information. Participants are given a list of 15 unrelated words, repeated over five different trials and are asked to repeat. Another list of 15 unrelated words are given and the client must again repeat the original list of 15 words and then again after 30 minutes.
\par
Several studies have shown that an impairment in RAVLT score reflect well the underlying pathology caused by AD. Thus making the RAVLT an effective early marker to detect AD in persons with memory complaints. Moradi investigated to what extent the RAVLT scores are predictable based on MRI data using machine learning approaches, as well as to find out what the most important brain regions are for the estimation of RAVLT scores. They found a highly significant cross validated correlation between the estimated and observed RAVLT immediate and RAVLT Percent Forgetting. Further, they found that the conversion of MCI subjects to AD in 3-years could be predicted based on either observed or estimated RAVLT scores with an accuracy comparible to MRI-based biomarkers \cite{Moradi2017}. Another study by Schoenberg found that RAVLT to best distinguish patients suspected of Alzheimer's disease from the psychiatric comparison group.
% Schoenberg - Test performance and classification statistics for the Rey Auditory Verbal Learning Test in selected clinical samples.

\subsection{Digit Symbol Substitution Test}
The Digit Symbol Substitution Test (DSST) involves a key consisting of the numbers 1-9 and a corresponding unique symbol. Below this key is a series of numbers from 1-9 in a randomized order and repeated multiple times. The participant is asked to fill in the corresponding symbol for each number. The task requires that the participant move between the key and the randomized sequence such that  they may retrieve the correct answer from the key, hold this in short-term memory and transcribe the key in the appropriate place.
% Rosano - Digit symbol substitution test and future clinical and subclinical disorders of cognition, mobility and mood in older adults.
Among those with no disorder in cognition, mobility and mood, being in the lowest DSST quartile compared to the highest was associated with nearly twice the odds of developing one or more clinical or subclinical disorders. Associations were stronger for incident clinical disorders in cognition. Slower psychomotor speed may serve as a biomarker of risk of clinical disorders, mobility and mood. While in part attributable to vascular brain disease, other potentially modifiable contributors may be present. Further studying the causes of psychomotor slowing with ageing might provide novel insights into age-related brain disorders. Pascoe compared patients with PD with Normal Cognition (PD-N) with those with PD and MCI (PD-MCI) and healthy participants. PD-MCI pariticipants achieved significantly lower scores than other groups in the DSST task.
% Pascoe - The symbol-digit modalities test in MCI: Evidence from Parkinson's Disease Patients.

\subsection{**Corsi Block Tapping Task}

@@

\subsection{Verbal Fluency}

@@

\subsection{Naming Tests}

@@

\subsection{Rey-Osterrieth Complex Figure Task}
% Cite Rey 1941 - L'examen psychologique dans les cas d'encephalopathie traumatique and Osterrieth 1944 - Le test de copie d'une figure complexe.
The Rey-Osterrieth Complex Figure (ROCF) is a task widely as a test of visuo-spatial skill and visual memory. The task, which was originally designed by Rey (1944) and standardised by Osterrieth (1944), asks a participant to copy a complex geometrical figure (known as the immediate copy condition and to recall and reproduce the figure from memory without warning (known as the delayed recall condition). In the immediate copy condition, the complexity of the figure requires an integrative cognitive ability. The reproduction of such a complex structure involves processes such as planning and organizational strategies that are related to executive functions.
% Salvadori - Qualitative Evaluation of the Immediate Copy of the Rey-Osterrieth Complex Figure: Comparison Between Vascular and Degenerative MCI patients.
Patients with vascular MCI had a worse performance in the immediate copy of the ROCF compared to individuals with degenerative MCI, despite their significant impairment in terms of general cognitive status and visual memory. Evidence shows that patients with disorders that possibly involve attention and executive functions are characterized by a more disorganised approach when copying the ROCF compared to controls.
% Teng - Persistence of Neuropsychological Testing Deficits in Mild Cognitive Impairment.

@@

\subsection{Hayling Sentence Completion Test}

@@

\subsection{Grooved Pegboard Test}

@@

\subsection{Visual Search}

@@

\subsection{Free Cued Selective Reminding Test}

@@

\section{Types of Language Assessment}

@@

\subsection{Interviews}

@@

\subsection{Picture-related content thematic elements}

@@

\subsection{**Perseveration}

@@

This section has looked at a number of different attributes which researchers have cited as being important in the detection of MCI and AD. Whilst in some cases there is a broad consensus about a given feature in the vast majority of cases there is some uncertainty. Some of the difficulty lies in different approaches used particularly around the criteria for experimental groups which means that it is difficult to be able to compare studies, like for like, as the populations of the participants being studied vary in subtle or obvious ways. It is important for any studies moving forward to use comparable standards when it comes to their experimental groups, such that the picture can be made clearer.

\section{**State of literature into Machine Learning and Natural Language processing techniques}
As we can see, diagnosing dementia through analysis has a large background in terms of psychological research. However, a lot of more current research has called for the use of machine learning as a way of assisting in the process of diagnosis \cite{Boschi2017}. Generally speaking, Natural Language Processing (NLP) systems mimic the semiotic perspective on language. That is to say that we can subdivide NLP processes into different components. This broadly fits with the literature review conducted by Emery \cite{Emery2000}.
\begin{enumerate}
	\item Morphological/Lexical: providing the basic language elements or vocabulary, such as words, their roots and inflections.
	\item Syntactic: for grouping and sequencing elements within samples of language (usually sentences).
	\item Semantic: for knowing the meaning of an utterance, usually defined in terms of the 'truth value' of the logical propositions that are thought to be expressed by sentences.
	\item Pragmatic: for understanding the context and purpose of an utterance.
\end{enumerate}
\subsubsection{Traditional methods of Machine Learning}
% Thomas - Automatic Detection and Rating of AD through the Lexical Analysis of Spontaneous Speech
Thomas et al, detail several lexical approaches to the problem of detecting and rating DAT. The approaches rely on character n-gram techniques but also explore correlation of usage of frequency of different types of speech. Their results act as a proof in concept of the utility of using a pure computational approach to the diagnosis of dementia using spontaneous speech. Results - 95\% accuracy dementia vs controls. 70\% accuracy classifying dementia into two categories. 50\% accuracy classifying dementia into four categories. They suggest further exploration of characteristics to classify.
% Bucks et al - Analysis of spontaneous, conversational speech in dementia of the Alzheimer's Type: Evaluation of an objective technique for analyzing lexical performance.
AD patients had higher mean P rate, A rate and V rate scores but lower N rate scores vs normal older controls \cite{Bucks2000}.
One of the first attempts to use machine learning and natural language techniques to look was conducted by Thomas \cite{Thomas2005} who was able to successfully demonstrate the ability of machine learning algorithms to analyse n-grams as well as other features to outperform a naive rule-based classifier which always selects the modal class. Orimaye et al (2017) \cite{Orimaye2017} investigated the use of machine learning algorithms to detect differences primarily in n-gram use to distinguish between those with a diagnosis of AD and healthy controls. Their main finding supported n-grams as the most significant predictor. One of the criticisms is the use of picture description tasks and n-grams. Because the language generated by this task is content specific the n-grams generated are only specific to the task given and cannot be generalised.
\subsubsection{The case for Deep Learning}

@@

The aim of my research is to ultimately find less burdensome ways of detecting dementia without the use of invasive procedures (taking bloods, or using medical equipment such as MRI's and EEG's) and without resorting to time-consuming and expensive psychological tests. There is a lot of research into the analysis of language as a marker for MCI \cite{Taler2008, Boschi2017} and there have been calls for research that explores the use of Natural Language Processing and Machine Learning techniques to aid in the analysis of language \cite{Boschi2017}. It is this area which is the focus of my research.

@@

I have deliberately selected the MCI population as examining language decline at this stage has the potential to further our understanding of how MCI progresses into dementia. Given that sampling a person's language is relatively effortless, my proposed research will look at whether we can find markers of MCI in natural language. I plan to explore the use Natural Language Processing and Machine Learning techniques to assist in this process. There have been a number of studies from the psychological perspective that look at the changes in language in those with Dementia. As documented above, these studies have identified a number of speech and language differences between healthy controls and those with dementia. Some of these features have been incorporated in work using machine learning to classify healthy controls from those with dementia. However, these have been done in lab settings, with formal cognitive assessment tasks such as the 'picture description task'. While this is helpful to learn that machine learning can be used to differentiate between healthy controls and those with dementia, this doesn't take the burden off healthcare professionals who have to conduct the assessments and other professionals who may be used to transcribe these assessments. Therefore we would like to see whether natural language, outside the framework of a deliberate language generating task, has the same ability to discriminate.

@@

Machine Learning itself is undergoing a shift in focus at the present time. I will describe this shift using the context of my proposed research. In traditional machine learning, the transcripts from cognitive assessments are used to generate abstract features such as idea density and more concrete features such as mean length of sentence or number of unique words. These features are then used to classify patients. As we have seen above, the picture with regards to what features deem important in this area is somewhat muddy. Machine Learning has a number of tools at it's disposal to aid this. The first is to statistically determine, given a model, which features contribute most highly to a given prediction.

@@

The broad objectives of my research are to explore whether the application of Natural Language Processing and Machine Learning techniques to spontaneous speech samples can improve our ability to diagnose Mild Cognitive Impairment in patients that will directly lead to Alzheimer's Disease. Another objective of my research is to provide some clarity as to which language features are most important to measure when looking for evidence of MCI and/or AD. Finally, an additional outcome of my research is the development of a repeatable battery of neuropsychological tests that specifically focus on being sensitive to Mild Cognitive Impairment.

@@

\section{**Research Challenges}
\subsection{Data Collection}

@@

\subsection{Co-morbidity of Language and Memory difficulties in MCI}Concerns: Language and Memory are quite naturally intertwined and it would be difficult to test one without some reliance on the other. I'm not going to control for memory problems as a potential confound, but does this weaken the research? How do I defend this? \newline
